---
title: "NMDS Reference Bugs- An example dataset"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
always_allow_html: yes
fig_width: 6 
fig_height: 4 
---

R version 4.0.3 (2020-10-10) -- "Bunny-Wunnies Freak Out"

Packages used for NMDS: vegan (version 2.5-7)

```{r setup, include=FALSE}


#R version 4.0.3 (2020-10-10) -- "Bunny-Wunnies Freak Out"
#Copyright (C) 2020 The R Foundation for Statistical Computing
#Platform: x86_64-w64-mingw32/x64 (64-bit)
#knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(dplyr)
library(lubridate)
library(purrr)
library(ggplot2)
library(readxl)
library(tidyr)

# Set working directory

setwd('C:/Backup/R/NMDS/NMDS')

```
## Methods


<font size="4"> The document serves as an example of analyses that will be conducted to identify natural variations in benthic communities across Virginia. These NMDS will support the Genus level IBI development process. This analysis is the first run of all of reference sites in Virginia. No West Virginia DEP data is used in this analysis. Several reference sites have been sent back to Virginia DEQ biologists for one more review. 


The dataset used is a subset of reference stations collected in Virginia. If stations appeared in the dataset more than 4 times, then the most recent 4 samples were used and the rest removed. Taxa that occurred in the dataset < 5% of the time were removed. The data was log10 +1 transformed. Environmental factors were compiled for each station and used to plot over the NMDS to show environmental variation associated with the community matrix. The envfit function in Vegan was used to plot the continuous environmental variables. </font>



```{r files, echo=FALSE}

###Load Data
bugs <- read_excel('benthics_ExcludedTaxaDropped_LBedits.xlsx', sheet='benthics_ExcludedTaxaDropped_LB')

bugs$Collection.Date<-as.Date(bugs$Collection.Date, format="%m/%d/%Y")

env <- read.csv('stationsSpatial_withAllData.csv')%>%
  subset(`X` == "Ref")

env$`Collection.Date`<-as.Date(env$`Collection.Date`, format="%m/%d/%Y")
#bugs<-read.csv("benthics_11June2021Final.csv")



Refbugs <-left_join(bugs,env, by=c("StationID", "Collection.Date"))%>%
  subset(`X`=="Ref")
Refbugs<-unique(Refbugs[ , 1:412])

refbugs<-write.csv("Refbugs", "data\\ReferenceBugs.csv", row.names=FALSE)
  





#DataSource, JRH_Final_Ref_Cod))
                            #, "Coastal", "StationID","UID",     #"Year","JulianDate","Collection.Date","Sta_Desc","Latitude", "Longitude","US_L3CODE",      #"US_L3NAME","US_L4CODE","US_L4NAME","VAHU6","ASSESS_REG", "OFFICE_NM","VaName", "Tidal",            # "VAHUSB", "FedName", "HUC10","Basin",            
#"BASIN_CODE", "Basin_Code", "CountyCityName", "FIPS", "BASINS_HUC_12",    
#"WQS_ID","WQS_CLASS", "WQS_SPSTDS","WQS_PWS", "WQS_TROUT",        
#"WQS_TIER_III", "Order",  "EDASOrder",  "totalArea_sqMile", "ELEVMIN",          
#"ELEVMAX", "ELEVMEAN", "ELEVSD",  "ELEVRANGE",  "SLPMIN",           
#"SLPMAX", "SLPMEAN", "SLPSD", "SLPRANGE", "wshdRain_mmyr",    
#"siteRain_mmyr", "wshdRain_inyr", "siteRain_inyr", "wshedRain_inyr", "BenSampID",        
#"RepNum", "Gradient", "Target.Count"))

#env<-read.csv("StationInfo_Final.csv")
#Bioregion<-read.csv("Bio_info.csv")
#stationinfo<-read.csv("Lucy_Example_StationInfo.csv")


###JRH used to check data
#Bugenv<-left_join(bugs,env, by="UID")

##Figure out how to join - removed collection date for now- LS had to remove bioregion

Bugenv<-left_join(env, Refbugs, by=c("StationID", "Collection.Date"))%>%
  select("StationID", "Year", "JulianDate", "Latitude", "Longitude", "Collection.Date", "Season" , "Gradient", "Coastal", "US_L3CODE", "US_L3NAME", "US_L4NAME", "ASSESS_REG", "Tidal", "VAHUSB", "Basin", "Basin_Code", "CountyCityName", "WQS_CLASS", "WQS_SPSTDS", "WQS_PWS", "WQS_TROUT", "WQS_TIER_III", "Order","EDASOrder", "totalArea_sqMile", "ELEVMEAN", "SLPMEAN", "wshdRain_mmyr", "siteRain_mmyr" )
#  unique()

#Bioregion2<-merge(x=Bioregion, y=stationinfo, by="StationID")%>%
#   select("StationID", "BioRegion", "Gradient")
# bugenv2<-merge(x=Bugenv, y=Bioregion2, by = "StationID")%>%
#   select("StationID", "Latitude", "Longitude", "WQM_STA_STRAHER_ORDER", "EPA_ECO_US_L3CODE","EPA_ECO_US_L3NAME",     #"BASINS_HUC_8_NAME","BASINS_VAHU6","WQS_SEC", "WQS_CLASS", "WQS_SPSTDS", "WQS_TROUT",
# "WQS_TIER_III", "Admin_Region", "Sta_Huc_Code", "Sta_Huc_Sbc_Bsc_Code", "Sta_Huc_Sbc_Code", "Sta_Wqm_Wat_Shed_Code", "Collection.Date", #"Season", "BioRegion", "Gradient")%>%
#   unique()

###Wait for lucy or emma here....###
#Bugenv$`Collection Date.x`<-as.Date(as.character(Bugenv$`Collection Date.x`), format="%m/%d/%Y")
#Bugenv$JulianDate<- yday(Bugenv$`Collection Date.x`)
Bugenv[sapply(Bugenv, is.character)] <- lapply(Bugenv[sapply(Bugenv, is.character)],as.factor) 


```

```{r remove rare taxa, echo=FALSE, message= FALSE}

#Emma's function to remove rare taxa

removeRareTaxa <- function(benthics, # wide tibble of taxa count by station
                           threshold, # numeric 
                           groupingFields # vector of all non taxa fields
){

  taxaMath <- benthics %>% 
    dplyr::select(-c( !!groupingFields) ) %>% 
    pivot_longer(cols = everything(), names_to = 'Taxa', values_to = 'Count') %>% 
    filter(!is.na(Count)) %>% 
    mutate(Count = ifelse(Count > 0, 1, 0)) %>% 
    group_by(Taxa) %>% 
    summarise(Count = sum(Count)) %>% 
    mutate(Percent = Count / nrow(benthics) * 100)
  # Separate these for easy reporting to user
  taxaToKeep <- filter(taxaMath, Percent >= !! threshold) 
  taxaToDrop <- filter(taxaMath, Percent < !! threshold) 
  
  benthicsSlim <- benthics %>% 
    group_by_at( groupingFields) %>% 
    pivot_longer(cols = -groupingFields, names_to = 'Taxa', values_to = 'Count') %>% 
    filter(Taxa %in% taxaToKeep$Taxa) %>% 
    replace(is.na(.), 0) %>% 
    mutate(Count = log1p(Count)) %>% 
    pivot_wider(names_from = Taxa, values_from = Count)
    
  return(
    list(`Taxa Math` = taxaMath,
         `Taxa Kept` = taxaToKeep,
         `Taxa Dropped` = taxaToDrop,
         `Benthics Slim` = benthicsSlim) )
}



dataList <- removeRareTaxa(Refbugs, # dataset to analyze, can filter on whatever you want beforehand (e.g. ref, stress, play with ecoregions)
                           threshold = 5, # drop taxa below 5% occurrence in dataset 
                           groupingFields = c( "StationID", "BenSampID.x" , "Collection DateTime", "RepNum.x", "Collection.Date")) # fields that you don't want thrown into the math
# how to view data that comes out of function

# how to view data that comes out of function
View(dataList$`Taxa Math`) # taxa occurrence and percent breakdown
View(dataList$`Taxa Kept`) # taxa higher or equal to input threshold
View(dataList$`Taxa Dropped`) # taxa lower than threshold, dropped from further analyses
View(dataList$`Benthics Slim`) # Taxa below threshold removed, individual counts logged


NMDSFive<-dataList$`Benthics Slim`

```



```{r pressure, echo=FALSE, message = FALSE}

#Removed and log transformed above! 

#bugs[is.na(bugs)]<-0


bugsnms<-NMDSFive%>%
  select("StationID", 6:119)

#already logged in JRH data
###buglog<-log1p(bugsnms[,2:112])


bugsnms$site<-1:nrow(bugsnms) 
###bugsnms$site<-1:nrow(bugsnms)


Site<-bugsnms%>%
 select("StationID","site")
BugNMS<- left_join(bugsnms,Site)%>%
  select(-c("site"))


```



<font size="4"> The first step was to read in the reference site bug taxa list and environmental factors dataset for each station. Join the environmental dataset with the bug dataset to account for multiple observations of each station and collection date and time. 

Check to make sure the bug and environmental join was successful: 

Number of rows in Community Matrix: `r format(nrow(Refbugs))`

Number or rows in Environmental Matrix: `r format(nrow(Bugenv))`

The data was log10+1 transformed. Rare taxa (<=5%) were removed. </font>









## Run NMDS for reference communities


```{r bugnms, echo=FALSE}
NMDSbug <-metaMDS (NMDSFive[,6:119],k=3, trymax=1000)

#Best run is 0.168
NMDSbug

#Stressplot fo NMDS
stressplot(NMDSbug)

plot(NMDSbug, "sites")
#orditorp(NMDSbug,display="sites",cex=1.25,air=0.01)


###Think Lucy was fitting genus to NMDS??
bug.spp.fit <- envfit(NMDSbug, bugsnms, permutations = 999) 
bug.spp.scrs <- as.data.frame(scores(bug.spp.fit, display = "vectors")) 


```

# Plot with Station IDs


```{r plot with stations, echo=FALSE, fig.width=12, fig.height=8}


samplescores<-as.data.frame(scores(NMDSbug))
samplescores$site<-Site$StationID

species.scores <- as.data.frame(scores(NMDSbug, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)

ggplot()+
  geom_text(data=samplescores,aes(x=NMDS1,y=NMDS2,label=site),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores,aes(x=NMDS1,y=NMDS2),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())




```

# Plot with Axis 3



```{r plot with axes 3, echo=FALSE, fig.width=12, fig.height=8}

ggplot()+
  geom_text(data=samplescores,aes(x=NMDS1,y=NMDS3,label=site),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores,aes(x=NMDS1,y=NMDS3),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

ggplot()+
  geom_text(data=samplescores,aes(x=NMDS2,y=NMDS3,label=site),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores,aes(x=NMDS2,y=NMDS3),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```


# Plot with Species


```{r plot with species, echo=FALSE, fig.width=12, fig.height=8}


#plot (all) with species
plot(NMDSbug, type = "n")
points(NMDSbug, display = "sites", cex = 0.8, pch=21, col="red", bg="yellow")
text(NMDSbug, display = "spec", cex=0.7, col="blue")


#samplescores<-as.data.frame(scores(NMDSbug))
#samplescores$site<-Site$StationID

#species.scores <- as.data.frame(scores(NMDSbug, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
#species.scores$species <- rownames(species.scores)

#ggplot()+
#  geom_text(data=samplescores,aes(x=NMDS1,y=NMDS2,label=site),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
#  geom_point(data=samplescores,aes(x=NMDS1,y=NMDS2),size=3)+
#  theme_bw() + 
#  theme(axis.text.x = element_blank(),  # remove x-axis text
#        axis.text.y = element_blank(), # remove y-axis text
#        axis.ticks = element_blank(),  # remove axis ticks
#        axis.title.x = element_text(size=18), # remove x-axis labels
#        axis.title.y = element_text(size=18), # remove y-axis labels
#        panel.background = element_blank(), 
#        panel.grid.major = element_blank(),  #remove major-grid labels
#        panel.grid.minor = element_blank(),  #remove minor-grid labels
#        plot.background = element_blank())




```


## Season 

Spring = Group 1
Fall = Group 2 

```{r plot Season, echo=FALSE, fig.width=12, fig.height=8}

par(mar = c(4, 4, .1, .1))

Bugenv$Season<- as.factor(Bugenv$Season)

#Season without taxa

plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, col=c("black", "red", "green"), pch = 16 [as.factor(as.numeric(Bugenv$Season))]) 
legend("topright", legend = c(levels(Bugenv$Season)),pch = 16, col=C("black", "red", "green"), cex=2)
legend("topleft", "stress = 0.16", bty = "n", cex = 2) # displays legend text of stress value 


#Season and bug NMDS
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, pch = c(16), col = colvec[Bugenv$Season]) 
legend("topright", legend = c(levels(Bugenv$Season)),pch = c( 16))
legend("topleft", "stress = 0.142", bty = "n", cex = 2) # displays legend text of stress value 
plot(bug.spp.fit, p.max = 0.001, col = "black", cex = 1) 

head(bug.spp.fit$vectors)

Season.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= Season2, distance="bray"))

Season.mrpp

```






## Ecoregion

```{r plot Ecoregion, echo=FALSE, fig.width=12, fig.height=8}

#Ecoregion

plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, pch=16, col= c("black","red", "green", "blue","purple","orange","yellow") [as.factor(as.numeric(Bugenv$US_L3NAME))]) 
legend("topright", legend = c(levels(Bugenv$US_L3NAME)),pch=16, col = c("black","red", "green", "blue","purple","orange","yellow"), bty = "n", cex = 1)
legend("topleft", "stress = 0.182", bty = "n", cex = 2) # displays legend text of stress value 


Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))

Ecoregion.mrpp

```

## Basin 

```{r plot Basin, echo=FALSE, fig.width=12, fig.height=8}

# Basin

plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, pch=16, col= c(unique(Bugenv$Basin_Code)) [as.factor(as.numeric(Bugenv$Basin_Code))]) 
legend("topright", legend = c(levels(Bugenv$Basin_Code)),pch=16, col = c(unique(Bugenv$Basin_Code)), bty = "n", cex = 1)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 


#Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))
Basin.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group=Basin_Code, distance="bray" ))

Basin.mrpp

```


## Admin Region

```{r plot Region, echo=FALSE, fig.width=12, fig.height=8}

#Admin-region
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites",cex=2, pch=16, col= c(unique(Bugenv$ASSESS_REG)) [as.factor(as.numeric(Bugenv$ASSESS_REG))]) 
legend("topright", legend = c(levels(Bugenv$ASSESS_REG)),pch=16, col = c(unique(Bugenv$ASSESS_REG)), bty = "n", cex = 2)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 


#Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))
admin.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group=ASSESS_REG, distance="bray" ))

admin.mrpp

```

## Sample Method
```{r plot Sample Method, echo=FALSE, fig.width=12, fig.height=8}

#Method
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, pch=16, col= c(unique(Bugenv$Gradient)) [as.factor(as.numeric(Bugenv$Gradient))]) 
legend("topright", legend = c(levels(Bugenv$Gradient)),pch=1, col = c(unique(Bugenv$Gradient)), bty = "n", cex = 2)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 

#Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))
Sample.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group=Gradient, distance="bray" ))

Sample.mrpp

```


## BioRegion
```{r plot Bioregion, echo=FALSE, fig.width=12, fig.height=8}

#Bioregion
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites",cex=2, pch=16,  col= c(unique(Bugenv$BioRegion)) [as.factor(as.numeric(Bugenv$BioRegion))]) 
legend("topright", legend = c(levels(Bugenv$BioRegion)),pch=16, col = c(unique(Bugenv$BioRegion)), bty = "n", cex = 2)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 

#Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))
Bioregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group=BioRegion, distance="bray" ))

Bioregion.mrpp

```


## BioRegion and Season 
```{r plot Bioregion and season, echo=FALSE, fig.width=12, fig.height=8}

#Bioregion and Season
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites",cex=2, pch=16,  col= c(unique(Bugenv$BioSeason)) [as.factor(as.numeric(Bugenv$BioSeason))]) 
legend("topright", legend = c(levels(Bugenv$BioSeason)),pch=16, col = c(unique(Bugenv$BioSeason)), bty = "n", cex = 2)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 

#Ecoregion.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group= US_L3NAME, distance="bray"))
BioSeason.mrpp<- with(Bugenv, mrpp(bugsnms[,2:126], group=BioSeason, distance="bray" ))

BioSeason.mrpp

```

## Stream Order - mising some stuff here, need more work, leaving out MRPP

```{r plot Stream order, echo=FALSE, fig.width=12, fig.height=8}

#Stream order
plot(NMDSbug, type = "n") #displays empty ordination space
points(NMDSbug, display = "sites", cex=2, pch=16, col= c(unique(Bugenv$EDASOrder)) [as.factor(as.numeric(Bugenv$EDASOrder))]) 
legend("topright", legend = c(levels(as.factor(as.numeric(Bugenv$EDASOrder)))),pch=16, col = c(unique(as.factor(as.numeric(Bugenv$EDASOrder)))), bty = "n", cex = 2)
legend("topleft", "stress = 0.18", bty = "n", cex = 2) # displays legend text of stress value 

#Stream.mrpp<- with(Bugenv, mrpp(BugNMS[,1:87], group=WQM_STA_STRAHER_ORDER, distance="bray" ))

#Stream.mrpp

```

