---
title: "Coastal NMS-No Midges"
output: html_document
---

R version 4.0.3 (2020-10-10) -- "Bunny-Wunnies Freak Out"

Packages used for NMDS: vegan (version 2.5-7)

```{r setup, include=FALSE}


#R version 4.0.3 (2020-10-10) -- "Bunny-Wunnies Freak Out"
#Copyright (C) 2020 The R Foundation for Statistical Computing
#Platform: x86_64-w64-mingw32/x64 (64-bit)
#knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(dplyr)
library(lubridate)
library(purrr) 
library(ggplot2)
library(readxl)
library(tidyr)
library(DT)
library(data.table)

# Set working directory

setwd('C:/Backup/R/NMDS/NMDS')

```
## Methods


<font size="4"> The document shows a series of NMDS ordinations for reference Coastal benthic communities in Virginia with environmental characteristics overlaid to evaluate natural differences in community compositions across coastal regions in Virginia. These NMDS will support the Genus level IBI development process. No West Virginia DEP data is used in this analysis.Reference sites were evaluated by regional biologists. 


The dataset used includes all coastal reference stations collected in Virginia that were deemed reference through a series or water quality parameter filters and regional biologist review. Coastal stations were considered to be located in the Southeastern Plains and MidAtlantic Coastal Plains ecoregion. Communities were deemed to have different biological communities based on ordinations conducted on the entire reference dataset as is demonstrated in the document titled "NMDS for Reference Streams". If stations appeared in the dataset more than 4 times, then the most recent 4 samples were used and the rest removed. Samples that had a total number of taxa below 100 collected at the time of sampling were also removed. Taxa that occurred in the dataset <= 5% of the time were removed. The data was log10 +1 transformed. Environmental factors were compiled for each station and used to plot over the NMDS to show environmental variation associated with the community matrix. The envfit function in Vegan was used to plot the continuous environmental variables. Some environmental variables like precipitation, slope, and elevation have not been calculated for all watersheds yet and will be added at a later date. </font>



```{r files, echo=FALSE}

###Load Data- don't need to do the environmental dataset unless things change
bugs <- read_excel('data//benthics_ExcludedTaxaDropped_NoMidges.xlsx', sheet='benthics_ExcludedTaxaDropped_LB')

bugs$Collection.Date<-as.Date(bugs$Collection.Date, format="%m/%d/%Y")

env <- read.csv('data//stationsSpatial_withAllData.csv')%>%
  subset(`X` == "Ref")

env$`Collection.Date`<-as.Date(env$`Collection.Date`, format="%m/%d/%Y")



# 
#  bioregion <- function(x) { 
#  if(x == 67 | x == 69 | x== 66) y <- "Mountain"
#  if(x  == 63 | x == 65 ) y <- "Coast"
#  if(x == 64 | x== 45) y <- "Piedmont"
#  return(y)
#  }
# # 
#  streamcategory<- function(x) { 
#  if(x == 1 | x == 2 ) y <- "Small"
#  if(x  == 3) y <- "Medium"
#  if(x == 4 | x== 5 | x== 6) y <- "Large"
#  return(y)
#  }
# # 
#  env$StreamCate<-sapply(env$Order, streamcategory)
# # 
#  env$Bioregion<-sapply(env$US_L3CODE, bioregion)
# 
 Env_coast<-env%>%
     subset(`US_L3NAME`=="Middle Atlantic Coastal Plain" | `US_L3NAME`==  "Southeastern Plains")%>%
   subset(Season!="Outside Sample Window")
 

   
Refbugs <-left_join(bugs,env, by=c("StationID", "Collection.Date"))%>%
  subset(`X`=="Ref")%>%
  subset(Season!= "Outside Sample Window")


coastal <-Refbugs%>%
  subset(`US_L3NAME` == "Middle Atlantic Coastal Plain"| `US_L3NAME`=="Southeastern Plains")


coastal<-unique(coastal[ , 1:411])

#refbugs_noncoast<-write.csv(Env_coast, "data\\Referenceenv_coast.csv", row.names=FALSE)



##opened saved spreadsheet and concatenated season, bioregion, and stream size results
#Pulled this dataset in for review 



Env_coast_2<-read.csv("data//Referenceenv_coast.csv")%>%
  subset(Season!="Outside Sample Window")

Env_coast_2$`Collection.Date`<-as.Date(Env_coast_2$`Collection.Date`, format="%m/%d/%Y")

##Figure out how to join - removed collection date for now- LS had to remove bioregion

Bugenv<-left_join(Env_coast_2, coastal, by=c("StationID", "Collection.Date"))%>%
  select("StationID", "Year", "JulianDate", "Latitude", "Longitude", "Collection.Date", "Season" , "Gradient", "Coastal", "US_L3CODE", "US_L3NAME", "US_L4NAME", "ASSESS_REG", "Tidal", "VAHUSB", "Basin", "Basin_Code", "CountyCityName", "WQS_CLASS", "WQS_SPSTDS", "WQS_PWS", "WQS_TROUT", "WQS_TIER_III", "Order","EDASOrder", "totalArea_sqMile", "ELEVMEAN", "SLPMEAN", "wshdRain_mmyr", "siteRain_mmyr", "Bioregion", "StreamCate", "Bioregionsize", "BioregionSeason" )
#  unique()


#Bugenv[sapply(Bugenv, is.character)] <- lapply(Bugenv[sapply(Bugenv, is.character)],as.factor) 


```

```{r remove rare taxa, echo=FALSE, message= FALSE}

#Emma's function to remove rare taxa

removeRareTaxa <- function(benthics, # wide tibble of taxa count by station
                           threshold, # numeric 
                           groupingFields # vector of all non taxa fields
){

  taxaMath <- benthics %>% 
    dplyr::select(-c( !!groupingFields) ) %>% 
    pivot_longer(cols = everything(), names_to = 'Taxa', values_to = 'Count') %>% 
    filter(!is.na(Count)) %>% 
    mutate(Count = ifelse(Count > 0, 1, 0)) %>% 
    group_by(Taxa) %>% 
    summarise(Count = sum(Count)) %>% 
    mutate(Percent = Count / nrow(benthics) * 100)
  # Separate these for easy reporting to user
  taxaToKeep <- filter(taxaMath, Percent >= !! threshold) 
  taxaToDrop <- filter(taxaMath, Percent < !! threshold) 
  
  benthicsSlim <- benthics %>% 
    group_by_at( groupingFields) %>% 
    pivot_longer(cols = -groupingFields, names_to = 'Taxa', values_to = 'Count') %>% 
    filter(Taxa %in% taxaToKeep$Taxa) %>% 
    replace(is.na(.), 0) %>% 
    mutate(Count = log1p(Count)) %>% 
    pivot_wider(names_from = Taxa, values_from = Count)
    
  return(
    list(`Taxa Math` = taxaMath,
         `Taxa Kept` = taxaToKeep,
         `Taxa Dropped` = taxaToDrop,
         `Benthics Slim` = benthicsSlim) )
}




datList_coast<-removeRareTaxa(coastal, # dataset to analyze, can filter on whatever you want beforehand (e.g. ref, stress, play with ecoregions)
                           threshold = 5, # drop taxa below 5% occurrence in dataset 
                           groupingFields = c( "StationID", "BenSampID.x" , "Collection DateTime", "RepNum.x", "Collection.Date")) # fields that you don't want thrown into the math
# how to view data that comes out of function

# how to view data that comes out of function
#View(dataList$`Taxa Math`) # taxa occurrence and percent breakdown
#View(dataList$`Taxa Kept`) # taxa higher or equal to input threshold
#View(dataList$`Taxa Dropped`) # taxa lower than threshold, dropped from further analyses
#View(dataList$`Benthics Slim`) # Taxa below threshold removed, individual counts logged


coastalFive<-datList_coast$`Benthics Slim`


```



```{r NMS, echo=FALSE, message = FALSE}


bugsnms_coast<-coastalFive%>%
  select("StationID", "Collection.Date", 6:105)


bugsnms_coast$site<-1:nrow(bugsnms_coast)


Site_coast<-bugsnms_coast%>%
 select("StationID","site","Collection.Date")
BugNMS_coast<- left_join(bugsnms_coast,Site_coast)%>%
  select(-c("site"))


```



<font size="4"> The first step was to read in the reference site bug taxa list and environmental factors dataset for each station. Join the environmental dataset with the bug dataset to account for multiple observations of each station and collection date and time. 

Check to make sure the bug and environmental join was successful: 

Number of rows in Community Matrix: `r format(nrow(Refbugs))`

Number or rows in Environmental Matrix: `r format(nrow(Bugenv))`

The data was log10+1 transformed. Rare taxa (<=5%) were removed. </font>


## NMDS Results for Coastal Streams

```{r bugnms noncoastal, echo=FALSE}

NMDSbug_coast <-metaMDS (coastalFive[,6:105],k=3, trymax=1000)

#Best run is 0.19
NMDSbug_coast

#Stressplot fo NMDS
stressplot(NMDSbug_coast)


#orditorp(NMDSbug,display="sites",cex=1.25,air=0.01)
plot(NMDSbug_coast, "sites")

samplescores_coast<-as.data.frame(scores(NMDSbug_coast))
samplescores_coast$StationID<-Site_coast$StationID
samplescores_coast$Collection.Date<- Site_coast$Collection.Date
samplescoresenv_coast<-left_join(samplescores_coast, Bugenv, by=c("StationID", "Collection.Date"))

env_continuous<-samplescoresenv_coast%>%
  select("StationID", "Year", "JulianDate", "Latitude", "Longitude", "totalArea_sqMile", "ELEVMEAN", "SLPMEAN", "wshdRain_mmyr", "siteRain_mmyr")




bug.sp.fit <- envfit(NMDSbug_coast, env_continuous[ , 2:9], permutations = 99, na.rm=TRUE) 
bug.sp.fit$vectors

coord_cont = as.data.frame(scores(bug.sp.fit, "vectors")) * ordiArrowMul(bug.sp.fit)
vectors<-coord_cont[-c(5,6), ]




```

## Plot with Stations: Coastal

```{r plot with stations noncoastal, echo=FALSE, fig.width=12, fig.height=8}


species.scores_coast <- as.data.frame(scores(NMDSbug_coast, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores_coast$species <- rownames(species.scores_coast)

ggplot()+
  geom_text(data=samplescores_coast,aes(x=NMDS1,y=NMDS2,label=StationID),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores_coast,aes(x=NMDS1,y=NMDS2),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())
ggplot()+
  geom_text(data=samplescores_coast,aes(x=NMDS2,y=NMDS3,label=StationID),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores_coast,aes(x=NMDS2,y=NMDS3),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


ggplot()+
  geom_text(data=samplescores_coast,aes(x=NMDS1,y=NMDS3,label=StationID),alpha=0.5, size=3,vjust=1.5,hjust=0.6) +  # add the species labels
  geom_point(data=samplescores_coast,aes(x=NMDS1,y=NMDS3),size=3)+
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


```


## Season: Coastal, No Midges

```{r plot Season noncoastal, echo=FALSE, fig.width=12, fig.height=8}
ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour =Season), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue", "black")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Season")



ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour =Season), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue", "black")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Season")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour =Season), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue", "black")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Season")


Season.mrpp.coastal<- mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$Season, distance="bray")

Season.mrpp.coastal

```




## Ecoregion: Coastal, No Midges

```{r plot Ecoregion noncoastal, echo=FALSE, fig.width=12, fig.height=8}

#Ecoregion

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = US_L3NAME), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size =6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "EcoRegion")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = US_L3NAME), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size =6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "EcoRegion")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = US_L3NAME), size = 5, alpha = 0.5) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size =6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "EcoRegion")



Ecoregion.mrpp.coast<- mrpp(bugsnms_coast[,6:105], group= samplescoresenv_coast$US_L3NAME, distance="bray")

Ecoregion.mrpp.coast

```


## Basin: Coastal, No Midges 

```{r plot Basin noncoastal, echo=FALSE, fig.width=12, fig.height=8}

# Basin
ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Basin_Code ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral", "cadetblue", "chartreuse", "cornflowerblue", "cornsilk4", "darkgrey", "darkorange2", "darkslateblue", "burlywood4", "darkgreen")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Basin")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Basin_Code ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral", "cadetblue", "chartreuse", "cornflowerblue", "cornsilk4", "darkgrey", "darkorange2", "darkslateblue", "burlywood4", "darkgreen")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Basin")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Basin_Code ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2", "coral", "cadetblue", "chartreuse", "cornflowerblue", "cornsilk4", "darkgrey", "darkorange2", "darkslateblue", "burlywood4", "darkgreen")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Basin")


Basin.mrpp.coast<- mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$Basin_Code, distance="bray" )

Basin.mrpp.coast

```


## Admin Region: Coastal, No Midges

```{r plot Region noncoastal, echo=FALSE, fig.width=12, fig.height=8}

#Admin-region

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = ASSESS_REG ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "DEQ Regional Office")


ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = ASSESS_REG ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "DEQ Regional Office")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = ASSESS_REG ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque", "darkorchid2")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "DEQ Regional Office")





admin.mrpp.coast<- mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$ASSESS_REG, distance="bray" )

admin.mrpp.coast

```



##Bioregion: Coastal, No Midges

```{r plot bioregion noncoastal, echo=FALSE, fig.width=12, fig.height=8}

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregion), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregion), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion")


ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregion), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion")


bioregion.mrpp.coast<- mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$Bioregion, distance="bray" )

bioregion.mrpp.coast


```


## Sample Method: Coastal, No Midges 

```{r plot Sample Method coastal, echo=FALSE, fig.width=12, fig.height=8}

#Method

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Gradient ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Sample Method")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Gradient ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Sample Method")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Gradient ), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
    annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Sample Method")



Sample.mrpp.coast<-mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$Gradient, distance="bray" )

Sample.mrpp.coast

```


## Stream Order: Coastal, No Midges

```{r plot Stream order coastal, echo=FALSE, fig.width=12, fig.height=8}

#Stream order
samplescoresenv_coast$Order<-as.character(samplescoresenv_coast$Order)

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Order), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Order), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Order), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")



Stream.mrpp.coast<-mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$Order, distance="bray" )

Stream.mrpp.coast

```

## Stream Order Categories: coastal, No Midges

```{r plot Stream order noncoastal category, echo=FALSE, fig.width=12, fig.height=8}

#Stream order


ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = StreamCate), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = StreamCate), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = StreamCate), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Order")



Cate.mrpp.coast<-mrpp(bugsnms_coast[,6:105], group=samplescoresenv_coast$StreamCate, distance="bray" )

Cate.mrpp.coast

```


## Water Quality Standard Class: Coastalx, No Midges

```{r plot Water quality Standard coastal, echo=FALSE, fig.width=12, fig.height=8}

#Water Quality Standard Class

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = WQS_CLASS), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Water Quality Standard")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = WQS_CLASS), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Water Quality Standard")

ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = WQS_CLASS), size = 5, alpha = 0.8) + 
     scale_colour_manual(values = c("blue", "deeppink", "cyan", "darkred", "bisque","darkorchid2" )) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Water Quality Standard")




Standard.mrpp.coast<- mrpp(bugsnms_coast[,6:105], samplescoresenv_coast$WQS_CLASS, distance="bray" )

Standard.mrpp.coast

```


## Natural Trout Water WQS: Coastal- No trout streams



## Bioregion and Season, No Midges

```{r plot bioregion and season noncoastal, echo=FALSE, fig.width=12, fig.height=8}


ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = BioregionSeason), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3",  "chocolate", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values = 16)+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion and Season")




ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = BioregionSeason), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3",  "chocolate", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values = 16)+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion and Season")



ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = BioregionSeason), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3",  "chocolate", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values = 16)+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Bioregion and Season")



basinseason.mrpp.coast<- mrpp(bugsnms_coast[,6:105], samplescoresenv_coast$BioregionSeason, distance="bray" )

basinseason.mrpp.coast

```



## Bioregion, Stream Size and Season: coastal, No Midges

```{r plot bioregion and season and stream size noncoastal, echo=FALSE, fig.width=12, fig.height=8}



ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregionsize), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3", "aquamarine3", "chocolate", "azure", "bisque","white","beige", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values = 16)+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream size, Bioregion, Season")

ggplot(data = samplescoresenv_coast, aes(x = NMDS2, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregionsize), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3", "aquamarine3", "chocolate", "azure", "bisque","white","beige", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values = c(0, 3, 16))+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream Size, Bioregion, Season")



ggplot(data = samplescoresenv_coast, aes(x = NMDS1, y = NMDS3)) + 
     geom_point(data = samplescoresenv_coast, aes(colour = Bioregionsize), size = 4, alpha = 0.8) + 
     scale_colour_manual(values = c("darkred","darkorchid2", "chartreuse", "cyan3", "aquamarine3", "chocolate", "azure", "bisque","white","beige", "deeppink", "blue", "darkmagenta", "cornflowerblue", "darkorange", "coral")) +
  scale_shape_manual(values =  16)+
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
       data = vectors, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = vectors, aes(x = NMDS1, y = NMDS2), colour = "grey30", 
       fontface = "bold", label = row.names(vectors)) + 
  annotate("text", x=0.5, y=1.3, label= "Stress= 0.19", size=6)+
     theme(axis.title = element_text(size = 14, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 15, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 15, colour = "grey30")) +
     labs(colour = "Stream, Bioregion, Season")


biosizeseason.mrpp.coast<- mrpp(bugsnms_coast[,6:105], samplescoresenv_coast$Bioregionsize, distance="bray" )

biosizeseason.mrpp.coast


```



```{r Summary table, echo=FALSE, fig.width=12, fig.height=8}

DT<- data.table(
  Grouping<- c("Season", "Ecoregion", "Basin", "Admin Region", "Bioregion", "Sample Method", "Stream Order", "Stream Order Category", "Water Quality Standards", "Trout", "Bioregion and season", "Bioregion, Size, season"), 
  `A Statistic`<- c(Season.mrpp.coastal$A, Ecoregion.mrpp.coast$A, Basin.mrpp.coast$A, admin.mrpp.coast$A, bioregion.mrpp.coast$A,Sample.mrpp.coast$A, Stream.mrpp.coast$A, Cate.mrpp.coast$A, Standard.mrpp.coast$A, "NA" ,basinseason.mrpp.coast$A, biosizeseason.mrpp.coast$A  )    , 
  `P-value`<- c(Season.mrpp.coastal$Pvalue, Ecoregion.mrpp.coast$Pvalue, Basin.mrpp.coast$Pvalue, admin.mrpp.coast$Pvalue, bioregion.mrpp.coast$Pvalue, Sample.mrpp.coast$Pvalue, Stream.mrpp.coast$Pvalue, Cate.mrpp.coast$Pvalue, Standard.mrpp.coast$Pvalue, "NA", basinseason.mrpp.coast$Pvalue, biosizeseason.mrpp.coast$Pvalue  )
)

DT<- format(DT, digits=3)

datatable(DT, class='cell-border stripe', rowname=FALSE, colnames=c("Grouping", "A Statisitc", "p-value"), caption="Table 1: A-statitics for MRPP for group parameters (999 permutations, method= Bray)", filter='top' )


```


